### 2023-07-05 chapters/interpretability/imli/experiment.tex
addressing comments

\begin{enumerate}
	\item How are the accuracy and size of classification rules generated by  {\imli} compared to existing interpretable classifiers?
	\item How is the scalability of {\imli} in solving large-scale classification problems compared to existing interpretable classifiers?
	\item How does {\imli} perform in terms of accuracy and scalability compared to classifiers that are non-interpretable? 
	\item How does the incremental learning in {\imli} perform compared to non-incremental MaxSAT-based learning in terms of accuracy, rule-sparsity, and scalability?
	\item How do different interpretable classification rules learned using {\imli} perform in terms of accuracy and rule-size?
	\item What are the effects of different hyper-parameters in {\imli}?
\end{enumerate}

\paragraph{Summary of Experimental Results.}
To summarize our experimental results, {\imli} achieves the best balance among prediction accuracy, interpretability, and scalability compared to existing interpretable rule-based classifiers. Particularly, compared to the most accurate classifier RIPPER, {\imli} demonstrates on average $ 1\% $ lower prediction accuracy, wherein the accuracy of {\imli} is higher than BRL, CORELS, and BRS in almost all datasets. In contrast, {\imli} generates significantly smaller rules than RIPPER, specifically in large datasets. Moreover, BRL, CORELS, and BRS report comparatively smaller rule-size than {\imli} on average, but with a significant decrease in accuracy.  In terms of scalability, {\imli} achieves the best performance compared to other interpretable and non-interpretable classifiers by classifying datasets with one million samples. While CORELS also scales to such large datasets,  its accuracy is lower than that of {\imli} by at least $ 2\% $ on average.  Therefore, {\imli} is not only scalable but also accurate in practical classification problems while also being interpretable by design.  We additionally analyze the comparative performance of different formulations presented in this chapter, where the  incremental approach empirically proves its efficiency than the na\"ive MaxSAT formulation.  Furthermore, we learn and compare the performance of different interpretable representations: decision lists, decision sets, CNF, and DNF formulas using {\imli} and present the efficacy of {\imli} in learning varied interpretable classifiers. Finally, we study the effect of different hyper-parameters in {\imli}, where each hyper-parameter provides a precise control among training time,  prediction accuracy, and rule-sparsity. In the following, we discuss our experimental results in detail.

@@

	\caption[Accuracy and rule-size of interpretable classifiers]{Comparison of accuracy and rule-size among interpretable classifiers. Each cell from the fourth to the eighth column contains test accuracy (top) and rule-size (bottom). `\textemdash' represents a timeout. Numbers in bold represent the best performing results among different classifiers.}
	\label{interpretability_imli_table:interpretable_classifiers} 
	%	\setlength{\tabcolsep}{.2em}
	\small{
	\begin{tabular}{lrrrrrrrrrrrr}

@@

\subsection{Experimental Results}
\label{interpretability_imli_sec:experiments_scalability}
\paragraph{Comparing {\imli} with interpretable classifiers.} We compare {\imli} with existing interpretable classifiers in three aspects: test accuracy, rule-size, and scalability.

\textit{Test accuracy and rule-size.}  We present the experimental  results of test accuracy and rule-size among interpretable classifiers in Table~\ref{interpretability_imli_table:interpretable_classifiers}, where the first, second, and third columns represent  the name of the dataset, the number of samples, and the number of features in the dataset, respectively. In each cell from the fourth to the eighth column in the table, the top value represents the median test accuracy and the bottom value represents the median size of rules measured through ten-fold cross-validation.

@@

We now compare {\imli} with each interpretable classifier in detail. Compared to RIPPER, {\imli} has lower accuracy in $ 9 $ out  of $ 12 $ datasets. More specifically, the accuracy of {\imli} is $ 1\% $ lower on average than RIPPER. The improved accuracy of RIPPER, however, results in the generation of higher size classification rules than {\imli} in most datasets. In particular, {\imli} generates sparser rules than RIPPER in $ 9 $ out of $ 12 $ datasets, wherein RIPPER times out in $ 3 $ datasets. Interestingly, the difference in rule-size is more significant in larger datasets, such as in `Vote' dataset, where RIPPER learns a classifier with $ 132 $ Boolean literals compared to $ 15 $ Boolean literals by {\imli}.  Therefore, {\imli} is better than RIPPER in terms of rule-sparsity, but lags slightly in accuracy.

{\imli} performs better than BRL both in terms of accuracy and rule-sparsity. In particular, {\imli} has  higher accuracy and lower rule-size than BRL in $ 12 $ and $ 8 $ datasets, respectively, in a total of $ 12 $ datasets, wherein BRL times out in $ 3 $ datasets.  While comparing with CORELS, {\imli} achieves higher accuracy in almost all datasets ($ 14 $ out of $ 15 $ datasets). A similar trend is observed in comparison with BRS, where {\imli} achieves higher accuracy in all of $ 12 $ datasets and BRS times out in $ 3 $ datasets. CORELS and BRS, however, generates sparser rules than {\imli} in most datasets, but by costing a significant decrease in accuracy.  For example, in the largest dataset `BNG(credit-g)' with $ 1 $ Million samples, BRS times out and CORELS generates a classifier with  $ 72.08\% $ accuracy with rule-size $ 2 $. {\imli}, in contrast, learns a classifier with $ 27.5 $ Boolean literals achieving $ 75.48\% $ accuracy, which is $ 3\% $ higher than CORELS. Therefore, {\imli}  makes a good balance between accuracy and rule-size compared to existing interpretable classifiers while also being highly scalable. In the following, we discuss the results on the scalability of all interpretable classifiers in detail.

@@

	\caption[Training time, test error, and rule-size of different formulations in {\imli}]{Comparison of training time, test error, and rule-size among different formulations presented in the chapter. In each cactus plot, the incremental formulation {\imli} with both mini-batch and iterative learning demonstrates the best performance in training time and test error than compared two formulations: non-incremental MaxSAT formulation {\imli}-$\mathsf{B}$ and incremental formulation with only mini-batch learning {\imli}-$\mathsf{M}$. In terms of rule-size, {\imli} often generates higher size rules than {\imli}-$\mathsf{M}$.}
	\label{interpretability_imli_fig:different_formulations}
\end{figure}

\paragraph{Comparison among different formulations in {\imli}.}  We compare the performance of different formulations for learning classification rules as presented in this chapter. In Figure~\ref{interpretability_imli_fig:different_formulations}, we show cactus plots for assessing training time (in seconds), test error (in percentage), and rule-size among different formulations. In the cactus plot, a point $ (x,y) $ denotes that the formulation yields lower than or equal to $ y $ training time (similarly, test error and rule-size) in $ x $ many instances in each dataset.

@@

Moving focus on the rule-size in the rightmost column in Figure~\ref{interpretability_imli_fig:different_formulations}, {\imli}-$\mathsf{B}$ achieves the highest rule-size in WDBC dataset. In contrast, the rule-size of {\imli}-$\mathsf{B}$  is lowest (zero) in Credit and Adult datasets. In the last two datasets, {\imli}-$\mathsf{B}$ times out during training and returns the default rule ``true'' by predicting all samples as class $ 1 $. The other two formulations {\imli} and {\imli}-$\mathsf{M}$  demonstrate a similar trend in rule-size in all datasets with {\imli}-$\mathsf{M}$ generating comparatively smaller size rules in Credit and Adult datasets. In this context,  the improvement of rule-sparsity of {\imli}-$\mathsf{M}$ is due to a comparatively higher test error (or lower accuracy) than {\imli} as observed in all three datasets. Therefore, {\imli} appears to be the best performing formulation w.r.t.\ training time, test error, and rule-size by balancing between accuracy and rule-size while being more scalable.

\begin{table*}
	\centering
	\caption[Accuracy and rule-size of different classification rules learned using {\imli}]{Comparison of test accuracy (top value) and rule-size (bottom value) among different rule-based representations learned using {\imli}. Numbers in bold denote the best performing results among different representations.}
	\label{interpretability_imli_table:different_representations}
	%	\setlength{\tabcolsep}{.2em}
	\small
	\begin{tabular}{lrrrrrrrrrrrrrrr}

@@

We deploy {\imli} to learn different interpretable rule-based representations: CNF and DNF classifiers, decision lists, and decision sets and present their comparative performance w.r.t.\ test accuracy and rule-size in Table~\ref{interpretability_imli_table:different_representations}. In each cell in this table, the top value represents the test accuracy and the bottom value represents the size of generated rules.

We learn all four interpretable representations on twelve datasets, where the CNF classifier appears to be the most accurate representation by achieving the highest accuracy in five datasets. In contrast, both DNF and decision lists achieve the highest accuracy in three datasets each; decision sets demonstrate the least performance in test accuracy by being more accurate in one dataset. To this end, the poor accuracy of decision sets is traded off by its rule-size as decision sets generate the sparsest rules compared to other representations. More precisely, decision sets have the smallest rule-size in six datasets, while CNF, DNF, and decision lists have the smallest rule-size in two, three, and one dataset, respectively. These results suggest that CNF classifiers are more favored in applications where higher accuracy is preferred, while decision sets are preferred in applications where higher interpretability is desired. In both cases, one could deploy {\imli} for learning varied representations of classification rules.

@@

	\caption[Effect of the number of clauses in {\imli}]{Effect of the number of clauses $ k $ on accuracy (test and train), rule-size, and training time. As $ k $ increases, both train and test accuracy of {\imli} increase while generating rules with higher size by incurring higher training time. }
	\label{interpretability_imli_fig:effect_of_number_of_clauses}
\end{figure}

\paragraph{Ablation study.} We experiment the effect of different hyper-parameters in {\imli} on prediction accuracy, rule-size, and training time in different datasets. In the following, we discuss the impact of the number of clauses, regularization parameter, and size of mini-batches in {\imli}.

@@

\textbf{Effect of regularizer $ \lambda $.} In Figure~\ref{interpretability_imli_fig:effect_of_number_of_regularization}, we vary $ \lambda $ in a logarithmic grid between $ 10^{-4} $ and $ 10^1 $. As stated in Eq.~\eqref{interpretability_imli_eq:obj_incr}, a higher value of $ \lambda $ puts more priority on the minimal changes in rules between consecutive mini-batches in incremental learning while allowing higher mini-batch errors. Thus, in the first and second columns in Figure~\ref{interpretability_imli_fig:effect_of_number_of_regularization}, as $ \lambda $ increases, both training and test accuracy gradually decrease. In addition, the size of rules (plots in the third column) also decreases. Finally, we observe that the training time generally decreases with $ \lambda $. This observation indicates that higher $ \lambda $ puts lower computational load to the MaxSAT solver as a fraction of training examples is allowed to be misclassified. Thus, similar to the number of clauses, regularization parameter $ \lambda $ in {\imli} allows to trade-off between accuracy and rule-size in a precise manner.

\textbf{Effect of the size of mini-batch.} In Figure~\ref{interpretability_imli_fig:effect_of_batch_size}, we present the effect of mini-batch size in {\imli}. As we consider more samples in a batch, both test and training accuracy increase in general as presented in the first and second columns in Figure~\ref{interpretability_imli_fig:effect_of_batch_size}. Similarly, the size of generated rules also increases with the number of samples. Due to solving higher size MaxSAT queries, the training time also increases in general with an increase in mini-batch size. Therefore, by varying the size of mini-batches, {\imli} allows controlling on training time vs the prediction accuracy (and rule-size) of generated rules.

@@

	\caption[Effect of regularization $ \lambda $ in {\imli}]{Effect of regularization $ \lambda $ on accuracy (test and train), rule-size, and training time. As $ \lambda $ increases, lower priority is given to accuracy. As a result, both training and test accuracy decrease with $ \lambda $ by generating smaller rules. }
	\label{interpretability_imli_fig:effect_of_number_of_regularization}
\end{figure}

@@

	\caption[Effect of batch-size in {\imli}]{Effect of bath size on accuracy (test and train), rule-size, and training time. As we consider more samples in a mini-batch, {\imli} generates more accurate and larger size classification rules.}
	\label{interpretability_imli_fig:effect_of_batch_size}
\end{figure}

>>>>

\begin{enumerate}
	\item How are the accuracy and size of classification rules generated by  {\imli} compared to existing interpretable classifiers?
	\item How is the scalability of {\imli} in solving large-scale classification problems compared to existing interpretable classifiers?
	\item How does {\imli} perform in terms of accuracy and scalability compared to classifiers that are non-interpretable? 
	\item How does the incremental learning in {\imli} perform compared to non-incremental MaxSAT-based learning in terms of accuracy, rule-sparsity, and scalability?
	\item How do different interpretable classification rules learned using {\imli} perform in terms of accuracy and rule size?
	\item What are the effects of different hyper-parameters in {\imli}?
\end{enumerate}

\paragraph{Summary of Experimental Results.}
To summarize our experimental results, {\imli} achieves the best balance among prediction accuracy, interpretability, and scalability compared to existing interpretable rule-based classifiers. Particularly, compared to the most accurate classifier RIPPER, {\imli} demonstrates on average $ 1\% $ lower prediction accuracy, wherein the accuracy of {\imli} is higher than BRL, CORELS, and BRS in almost all datasets. In contrast, {\imli} generates significantly smaller rules than RIPPER, specifically in large datasets. Moreover, BRL, CORELS, and BRS report comparatively smaller rule size than {\imli} on average, but with a significant decrease in accuracy.  In terms of scalability, {\imli} achieves the best performance compared to other interpretable and non-interpretable classifiers by classifying datasets with one million samples. While CORELS also scales to such large datasets,  its accuracy is lower than that of {\imli} by at least $ 2\% $ on average.  Therefore, {\imli} is not only scalable but also accurate in practical classification problems while also being interpretable by design.  We additionally analyze the comparative performance of different formulations presented in this chapter, where the  incremental approach empirically proves its efficiency than the na\"ive MaxSAT formulation.  Furthermore, we learn and compare the performance of different interpretable representations: decision lists, decision sets, CNF, and DNF formulas using {\imli} and present the efficacy of {\imli} in learning varied interpretable classifiers. Finally, we study the effect of different hyper-parameters in {\imli}, where each hyper-parameter provides a precise control among training time,  prediction accuracy, and rule-sparsity. In the following, we discuss our experimental results in detail.

@@

	\caption[Accuracy and rule-size of interpretable classifiers]{Comparison of accuracy and rule size among interpretable classifiers. Each cell from the fourth to the eighth column contains test accuracy (top) and rule size (bottom). `\textemdash' represents a timeout. Numbers in bold represent the best performing results among different classifiers.}
	\label{interpretability_imli_table:interpretable_classifiers} 
	%	\setlength{\tabcolsep}{.2em}
	\small{
	\begin{tabular}{lrrrrrrrrrrrr}

@@

\subsection{Experimental Results}
\label{interpretability_imli_sec:experiments_scalability}
\paragraph{Comparing {\imli} with interpretable classifiers.} We compare {\imli} with existing interpretable classifiers in three aspects: test accuracy, rule size, and scalability.

\textit{Test accuracy and rule size.}  We present the experimental  results of test accuracy and rule size among interpretable classifiers in Table~\ref{interpretability_imli_table:interpretable_classifiers}, where the first, second, and third columns represent  the name of the dataset, the number of samples, and the number of features in the dataset, respectively. In each cell from the fourth to the eighth column in the table, the top value represents the median test accuracy and the bottom value represents the median size of rules measured through ten-fold cross-validation.

@@

We now compare {\imli} with each interpretable classifier in detail. Compared to RIPPER, {\imli} has lower accuracy in $ 9 $ out  of $ 12 $ datasets. More specifically, the accuracy of {\imli} is $ 1\% $ lower on average than RIPPER. The improved accuracy of RIPPER, however, results in the generation of higher size classification rules than {\imli} in most datasets. In particular, {\imli} generates sparser rules than RIPPER in $ 9 $ out of $ 12 $ datasets, wherein RIPPER times out in $ 3 $ datasets. Interestingly, the difference in rule size is more significant in larger datasets, such as in `Vote' dataset, where RIPPER learns a classifier with $ 132 $ Boolean literals compared to $ 15 $ Boolean literals by {\imli}.  Therefore, {\imli} is better than RIPPER in terms of rule-sparsity, but lags slightly in accuracy.

{\imli} performs better than BRL both in terms of accuracy and rule-sparsity. In particular, {\imli} has  higher accuracy and lower rule size than BRL in $ 12 $ and $ 8 $ datasets, respectively, in a total of $ 12 $ datasets, wherein BRL times out in $ 3 $ datasets.  While comparing with CORELS, {\imli} achieves higher accuracy in almost all datasets ($ 14 $ out of $ 15 $ datasets). A similar trend is observed in comparison with BRS, where {\imli} achieves higher accuracy in all of $ 12 $ datasets and BRS times out in $ 3 $ datasets. CORELS and BRS, however, generates sparser rules than {\imli} in most datasets, but by costing a significant decrease in accuracy.  For example, in the largest dataset `BNG(credit-g)' with $ 1 $ Million samples, BRS times out and CORELS generates a classifier with  $ 72.08\% $ accuracy with rule size $ 2 $. {\imli}, in contrast, learns a classifier with $ 27.5 $ Boolean literals achieving $ 75.48\% $ accuracy, which is $ 3\% $ higher than CORELS. Therefore, {\imli}  makes a good balance between accuracy and rule size compared to existing interpretable classifiers while also being highly scalable. In the following, we discuss the results on the scalability of all interpretable classifiers in detail.

@@

	\caption[Training time, test error, and rule size of different formulations in {\imli}]{Comparison of training time, test error, and rule size among different formulations presented in the chapter. In each cactus plot, the incremental formulation {\imli} with both mini-batch and iterative learning demonstrates the best performance in training time and test error than compared two formulations: non-incremental MaxSAT formulation {\imli}-$\mathsf{B}$ and incremental formulation with only mini-batch learning {\imli}-$\mathsf{M}$. In terms of rule size, {\imli} often generates higher size rules than {\imli}-$\mathsf{M}$.}
	\label{interpretability_imli_fig:different_formulations}
\end{figure}

\paragraph{Comparison among different formulations in {\imli}.}  We compare the performance of different formulations for learning classification rules as presented in this chapter. In Figure~\ref{interpretability_imli_fig:different_formulations}, we show cactus plots for assessing training time (in seconds), test error (in percentage), and rule size among different formulations. In the cactus plot, a point $ (x,y) $ denotes that the formulation yields lower than or equal to $ y $ training time (similarly, test error and rule size) in $ x $ many instances in each dataset.

@@

Moving focus on the rule size in the rightmost column in Figure~\ref{interpretability_imli_fig:different_formulations}, {\imli}-$\mathsf{B}$ achieves the highest rule size in WDBC dataset. In contrast, the rule size of {\imli}-$\mathsf{B}$  is lowest (zero) in Credit and Adult datasets. In the last two datasets, {\imli}-$\mathsf{B}$ times out during training and returns the default rule ``true'' by predicting all samples as class $ 1 $. The other two formulations {\imli} and {\imli}-$\mathsf{M}$  demonstrate a similar trend in rule size in all datasets with {\imli}-$\mathsf{M}$ generating comparatively smaller size rules in Credit and Adult datasets. In this context,  the improvement of rule-sparsity of {\imli}-$\mathsf{M}$ is due to a comparatively higher test error (or lower accuracy) than {\imli} as observed in all three datasets. Therefore, {\imli} appears to be the best performing formulation w.r.t.\ training time, test error, and rule size by balancing between accuracy and rule size while being more scalable.

\begin{table*}
	\centering
	\caption[Accuracy and rule-size of different classification rules learned using {\imli}]{Comparison of test accuracy (top value) and rule size (bottom value) among different rule-based representations learned using {\imli}. Numbers in bold denote the best performing results among different representations.}
	\label{interpretability_imli_table:different_representations}
	%	\setlength{\tabcolsep}{.2em}
	\small
	\begin{tabular}{lrrrrrrrrrrrrrrr}

@@

We deploy {\imli} to learn different interpretable rule-based representations: CNF and DNF classifiers, decision lists, and decision sets and present their comparative performance w.r.t.\ test accuracy and rule size in Table~\ref{interpretability_imli_table:different_representations}. In each cell in this table, the top value represents the test accuracy and the bottom value represents the size of generated rules.

We learn all four interpretable representations on twelve datasets, where the CNF classifier appears to be the most accurate representation by achieving the highest accuracy in five datasets. In contrast, both DNF and decision lists achieve the highest accuracy in three datasets each; decision sets demonstrate the least performance in test accuracy by being more accurate in one dataset. To this end, the poor accuracy of decision sets is traded off by its rule size as decision sets generate the sparsest rules compared to other representations. More precisely, decision sets have the smallest rule size in six datasets, while CNF, DNF, and decision lists have the smallest rule size in two, three, and one dataset, respectively. These results suggest that CNF classifiers are more favored in applications where higher accuracy is preferred, while decision sets are preferred in applications where higher interpretability is desired. In both cases, one could deploy {\imli} for learning varied representations of classification rules.

@@

	\caption[Effect of the number of clauses in {\imli}]{Effect of the number of clauses $ k $ on accuracy (test and train), rule size, and training time. As $ k $ increases, both train and test accuracy of {\imli} increase while generating rules with higher size by incurring higher training time. }
	\label{interpretability_imli_fig:effect_of_number_of_clauses}
\end{figure}

\paragraph{Ablation study.} We experiment the effect of different hyper-parameters in {\imli} on prediction accuracy, rule size, and training time in different datasets. In the following, we discuss the impact of the number of clauses, regularization parameter, and size of mini-batches in {\imli}.

@@

\textbf{Effect of regularizer $ \lambda $.} In Figure~\ref{interpretability_imli_fig:effect_of_number_of_regularization}, we vary $ \lambda $ in a logarithmic grid between $ 10^{-4} $ and $ 10^1 $. As stated in Eq.~\eqref{interpretability_imli_eq:obj_incr}, a higher value of $ \lambda $ puts more priority on the minimal changes in rules between consecutive mini-batches in incremental learning while allowing higher mini-batch errors. Thus, in the first and second columns in Figure~\ref{interpretability_imli_fig:effect_of_number_of_regularization}, as $ \lambda $ increases, both training and test accuracy gradually decrease. In addition, the size of rules (plots in the third column) also decreases. Finally, we observe that the training time generally decreases with $ \lambda $. This observation indicates that higher $ \lambda $ puts lower computational load to the MaxSAT solver as a fraction of training examples is allowed to be misclassified. Thus, similar to the number of clauses, regularization parameter $ \lambda $ in {\imli} allows to trade-off between accuracy and rule size in a precise manner.

\textbf{Effect of the size of mini-batch.} In Figure~\ref{interpretability_imli_fig:effect_of_batch_size}, we present the effect of mini-batch size in {\imli}. As we consider more samples in a batch, both test and training accuracy increase in general as presented in the first and second columns in Figure~\ref{interpretability_imli_fig:effect_of_batch_size}. Similarly, the size of generated rules also increases with the number of samples. Due to solving higher size MaxSAT queries, the training time also increases in general with an increase in mini-batch size. Therefore, by varying the size of mini-batches, {\imli} allows controlling on training time vs the prediction accuracy (and rule size) of generated rules.

@@

	\caption[Effect of regularization $ \lambda $ in {\imli}]{Effect of regularization $ \lambda $ on accuracy (test and train), rule size, and training time. As $ \lambda $ increases, lower priority is given to accuracy. As a result, both training and test accuracy decrease with $ \lambda $ by generating smaller rules. }
	\label{interpretability_imli_fig:effect_of_number_of_regularization}
\end{figure}

@@

	\caption[Effect of batch-size in {\imli}]{Effect of bath size on accuracy (test and train), rule size, and training time. As we consider more samples in a mini-batch, {\imli} generates more accurate and larger size classification rules.}
	\label{interpretability_imli_fig:effect_of_batch_size}
\end{figure}

### 2023-07-05 chapters/interpretability/imli/introduction.tex
addressing comments

The problem of learning rule-based classifiers is known to be computationally intractable. The earliest tractable approaches for classifiers such as decision trees and decision lists relied on heuristically chosen objective functions and greedy algorithmic techniques~\cite{ClarkN1989,CohenS1999,quinlan2014}. In these approaches, the size of rules is controlled by early stopping,  ad-hoc rule pruning, etc. In recent approaches, the interpretable classification problem is reduced to an optimization problem, where the accuracy and the sparsity of rules are optimized jointly~\cite{lakkaraju2016interpretable,narodytska2018learning}. Different optimization solvers such as linear programming~\cite{malioutov2013exact}, sub-modular optimizations~\cite{lakkaraju2016interpretable}, and Bayesian optimizations~\cite{letham2015interpretable} are then deployed to find the best classifier with maximum accuracy and minimum rule-size. In our study, we study an alternate optimization approach that fits particularly well to rule-learning problems. Particularly,  we study and improve a \textit{maximum satisfiability} (MaxSAT)  solution for learning interpretable rule-based classifiers. Building on MaxSAT, ~\cite{malioutov2018mlic} considers a learning framework to jointly optimize the accuracy and the size of classification rules. The said approach constructs and solves a MaxSAT query of $ \mathcal{O}(kn) $  size (number of clauses) to learn a $ k $-clause CNF formula on a dataset containing $ n $ samples (Chapter ~\ref{interpretability_imli_sec:encoding}). Naturally, the na\"ive formulation cannot scale to large values of $ n $ and $ k $. To scale learning to large datasets, we discuss an incremental learning technique along with the MaxSAT formulation.

>>>>

The problem of learning rule-based classifiers is known to be computationally intractable. The earliest tractable approaches for classifiers such as decision trees and decision lists relied on heuristically chosen objective functions and greedy algorithmic techniques~\cite{ClarkN1989,CohenS1999,quinlan2014}. In these approaches, the size of rules is controlled by early stopping,  ad-hoc rule pruning, etc. In recent approaches, the interpretable classification problem is reduced to an optimization problem, where the accuracy and the sparsity of rules are optimized jointly~\cite{lakkaraju2016interpretable,narodytska2018learning}. Different optimization solvers such as linear programming~\cite{malioutov2013exact}, sub-modular optimizations~\cite{lakkaraju2016interpretable}, and Bayesian optimizations~\cite{letham2015interpretable} are then deployed to find the best classifier with maximum accuracy and minimum rule size. In our study, we study an alternate optimization approach that fits particularly well to rule-learning problems. Particularly,  we study and improve a \textit{maximum satisfiability} (MaxSAT)  solution for learning interpretable rule-based classifiers. Building on MaxSAT, ~\cite{malioutov2018mlic} considers a learning framework to jointly optimize the accuracy and the size of classification rules. The said approach constructs and solves a MaxSAT query of $ \mathcal{O}(kn) $  size (number of clauses) to learn a $ k $-clause CNF formula on a dataset containing $ n $ samples (Chapter ~\ref{interpretability_imli_sec:encoding}). Naturally, the na\"ive formulation cannot scale to large values of $ n $ and $ k $. To scale learning to large datasets, we discuss an incremental learning technique along with the MaxSAT formulation.

### 2023-07-05 chapters/interpretability/imli/problem_formulation.tex
addressing Kuldeep's comments
addressing comments

 We learn classifiers that balance two 
 goals: of  being accurate but also interpretable.  
 Various notions of interpretability have been proposed in the context of   classification problems. A common proxy for interpretability in the context of decision rules 
 is the sparsity of rule. For instance, a rule involving fewer literals is highly interpretable.  In this chapter, we minimize the total number of literals in all clauses, which motivates us to  find $ \Rule  $ with minimum  $ |\Rule| $. Let $ \Rule $ classify all samples correctly during training. Among all the classification rules that classify all samples correctly,  we choose the sparsest (most interpretable) such $ \Rule $.

@@

Higher  values of $ \lambda $ generate a rule with a smaller rule-size but of more training errors, and vice-versa. Thus, $ \lambda $ can be tuned to trade-off between accuracy and interpretability for a rule-based classifier, which we experiment extensively in Chapter~\ref{interpretability_imli_sec:experiments}.

>>>>

 We learn classifiers that balance two 
 goals: of  being accurate but also interpretable.  
 Various notions of interpretability have been discussed in the context of   classification problems. A common proxy for interpretability in the context of decision rules 
 is the sparsity of rule. For instance, a rule involving fewer literals is highly interpretable.  In this chapter, we minimize the total number of literals in all clauses, which motivates us to  find $ \Rule  $ with minimum  $ |\Rule| $. Let $ \Rule $ classify all samples correctly during training. Among all the classification rules that classify all samples correctly,  we choose the sparsest (most interpretable) such $ \Rule $.

@@

Higher  values of $ \lambda $ generate a rule with a smaller rule size but of more training errors, and vice-versa. Thus, $ \lambda $ can be tuned to trade-off between accuracy and interpretability for a rule-based classifier, which we experiment extensively in Chapter~\ref{interpretability_imli_sec:experiments}.

### 2023-07-05 chapters/interpretability/imli/related_work.tex
addressing comments

Recently, several optimization frameworks have been proposed for interpretable classification, where both accuracy and rule-size are optimized during training. For example,~\cite{malioutov2013exact} proposed exact learning of rule-based classifiers based on Boolean compressed sensing using a linear programming formulation.~\cite{su2016learning} presented two-level Boolean rules, where the trade-off between classification accuracy and interpretability is studied. In their work, the Hamming loss is used to characterize prediction accuracy, and sparsity is used to characterize the interpretability of rules.~\cite{wang2015falling} proposed a Bayesian optimization framework for learning falling rule lists, which is an ordered list of if-then rules. Other similar approaches based on Bayesian analysis for learning classification rules are~\cite{letham2015interpretable,wang2017bayesian}. Building on custom discrete optimization techniques,~\cite{angelino2017learning} proposed an optimal learning technique for decision lists using a branch-and-bound algorithm. In a separate study,~\cite{lakkaraju2016interpretable} highlighted the importance of decision sets over decision lists in terms of interpretability and considered a sub-modular optimization problem for learning a near-optimal solution for decision sets. Our approach for interpretable classification, however, relies on the improvement in formal methods over the decades, particularly the efficient CDCL-based solution for satisfiability (SAT) problems~\cite{silva2003grasp}.

>>>>

Recently, several optimization frameworks have been proposed for interpretable classification, where both accuracy and rule size are optimized during training. For example,~\cite{malioutov2013exact} proposed exact learning of rule-based classifiers based on Boolean compressed sensing using a linear programming formulation.~\cite{su2016learning} presented two-level Boolean rules, where the trade-off between classification accuracy and interpretability is studied. In their work, the Hamming loss is used to characterize prediction accuracy, and sparsity is used to characterize the interpretability of rules.~\cite{wang2015falling} proposed a Bayesian optimization framework for learning falling rule lists, which is an ordered list of if-then rules. Other similar approaches based on Bayesian analysis for learning classification rules are~\cite{letham2015interpretable,wang2017bayesian}. Building on custom discrete optimization techniques,~\cite{angelino2017learning} proposed an optimal learning technique for decision lists using a branch-and-bound algorithm. In a separate study,~\cite{lakkaraju2016interpretable} highlighted the importance of decision sets over decision lists in terms of interpretability and considered a sub-modular optimization problem for learning a near-optimal solution for decision sets. Our approach for interpretable classification, however, relies on the improvement in formal methods over the decades, particularly the efficient CDCL-based solution for satisfiability (SAT) problems~\cite{silva2003grasp}.
